\documentclass{article}
\usepackage[mathletters]{ucs}
\usepackage[utf8x]{inputenc}
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}


\setcounter{secnumdepth}{0}
\title{With Software As a Service, Is Only the Network Luddite Free?}
\author{Bradley M. Kuhn}
\date{Sunday 1 May 2011}
\begin{document}
\maketitle

\section{In The Beginning \ldots{}}

\begin{itemize}
\item 
  There was a computer.

\item 
  There were terminals.

\item 
  There were users.

\item 
  \ldots{} and the users had freedom.

\end{itemize}
\section{Ye Ol' Four Freedoms}

\begin{itemize}
\item 
  to use.

\item 
  to learn and modify for yourself.

\item 
  to copy and share.

\item 
  to modify and share modified versions.

\end{itemize}
\section{… Then Freedom Eluded Us}

\begin{itemize}
\item 
  They built licensing walls.

\item 
  By separating the software from the computer.

\end{itemize}
\begin{itemize}
\item 
  Who Invented Licensing?
\end{itemize}
\section{Who Invented Licensing?}

Some think he did:
Hai Hobbyists! Softwearz gived free-domes, but I ated them.
\begin{itemize}
\item 
  Gates convinces IBM to license rather then acquire DOS.
\end{itemize}
\section{Who Fixed this Problem?}

\section{Who Fixed this Problem?}

\section{RMS}

\begin{itemize}
\item 
  Job at MIT's AI lab in late 1970s and early 1980s.

\item 
  Programming their PDP--10!

\item 
  In the Golden Age.

\end{itemize}
\section{MIT AI Lab: The Golden Age}

\begin{itemize}
\item 
  academic computing.
\item 
  software sharing.
\item 
  no licenses!
\end{itemize}
\section{MIT AI Lab}

MIT changed.

\section{MIT AI Lab}

patents == \$\$\$

\section{MIT AI Lab}

spin-offs == \$\$\$

\section{GNU}

\begin{itemize}
\item 
  Changes the world.

\item 
  For the user of individual computers.

\item 
  No reason now not to have freedom on your own machine.

\end{itemize}
\section{Internet changes little (to start)}

\begin{itemize}
\item 
  Client/Server computing.

\item 
  Freedom implications are basically the same.

\item 
  Example: The Email System.

\end{itemize}
\section{Email: The MTA}

\begin{itemize}
\item 
  Yours has freedom: (postfix, exim, sendmail).

\item 
  Plenty of proprietary ones: (exchange).

\item 
  RFCs define interoperability.

\item 
  We reverse-engineer RFC-less protocols: (OpenChange).

\end{itemize}
\section{Email: The MUA}

\begin{itemize}
\item 
  Yours has freedom: (Thunderbird, mutt, Gnus)

\item 
  Plenty of proprietary ones: (outlook)

\end{itemize}
\section{The Great Fear: SaaS}

\begin{itemize}
\item 
  SaaS: Software as a Service

\item 
  True “SaaS” would end software freedom.

\item 
  Imagine: GCC as an network service.

\item 
  SaaS == your private computing done elsewhere.

\item 
  Copyeft proponents all feared this threat.

\item 
  Is not widspread: but needs constant vigilance:
  \begin{itemize}
  \item 
    http://www.gnu.org/philosophy/who-does-that-server-really-serve.html
  \end{itemize}
\end{itemize}
\section{A New Era of Computing?}

Anyone know what this is?

☁

\section{New Era: Cloud Computing?}

\begin{itemize}
\item 
  Unicode Character `CLOUD' (U+2601)
\item 
  Given what this looks like …
  \begin{itemize}
  \item 
    … Unicode committee was omniscient.
  \end{itemize}
\end{itemize}
☁
\section{Cloud Computing: What changed?}

\begin{itemize}
\item 
  The Browser as application delivery platform.

\item 
  AJAX.

\item 
  Whose computing is done where?

\item 
  Could Computing: more complicated than SaaS.

\item 
  Interesting applications are collaboration.

\end{itemize}
\section{Email becomes Gmail}

\begin{itemize}
\item 
  ``The experience''

\item 
  Effectively: Thin Client.

\item 
  Mixes Up Free again: Price vs. Freedom.

\item 
  RFCs no longer enough.

\item 
  We're back to proprietary ``lock-in''.

\item 
  Users like it for collaboration aspects.

\end{itemize}
\section{Users and “The Cloud”}

\section{Users and “The Cloud”}

Even \emph{Worse} Than Proprietary Apps and OS'es
\section{What's The Challenge?}

\begin{itemize}
\item 
  Not merely a question of code and licensing.
  \begin{itemize}
  \item 
    Although that's part of it (more later).
  \end{itemize}
\item 
  A question of community source+data control.

\item 
  Power to transport: applications, data, human connections.

\item 
  Autonomy: of code, of data, of community.

\end{itemize}
\section{How Do We Start?}

\begin{itemize}
\item 
  Infrastructure that respects freedom.

\item 
  Empower the user: give them \textbf{all} the code.

\end{itemize}
\section{Can't Stop There}

\begin{itemize}
\item 
  User communities need ability to transplant.

\item 
  Reclaim your data; relocate your community.

\item 
  Truly distributed data.

\item 
  This is \textbf{tough} programming.

\end{itemize}
\section{Projects to Look At}

\begin{itemize}
\item 
  status.net identi.ca (laconi.ca).
  \begin{itemize}
  \item 
    DTRT in this regard for Microblogging
  \item 
    Follow me there: `bkuhn'
  \end{itemize}
\item 
  Jesse Vincent's Prophet, CouchDB, and other truly decentralized
  data stores.

\item 
  Yourselves.
  \begin{itemize}
  \item 
    make it even easier for developers to respect freedom.
  \end{itemize}
\end{itemize}
\section{License To Look At}

\begin{itemize}
\item 
  GPL is the BSD of Network Service.

\item 
  Affero GPL.

\item 
  Extends copyleft to network service world.

\item 
  Handles the code side well.

\end{itemize}
\section{Techniques to Look At}

\begin{itemize}
\item 
  Deployed applications auto-give users source.

\item 
  Data is downloadable in community-chunks.

\end{itemize}
\section{``Future'' of the Cloud}

\begin{itemize}
\item 
  Disjoint but Integrated.

\item 
  Portable and Autonomous.

\item 
  Developers decide next direction.

\item 
  Ask \textbf{the} question in Cloud projects:
  \begin{itemize}
  \item 
    Are we respecting the users' freedom and autonomy?
  \end{itemize}
\item 
  The data is theirs; we are \emph{their} custodians.

\end{itemize}
\section{More Info and This Talk's License}

\begin{itemize}
\item 
  More Info?
  \begin{itemize}
  \item 
    http://www.gnu.org/philosophy/who-does-that-server-really-serve.html
  \item 
    http://autonomo.us/
  \item 
    http://status.net/
  \end{itemize}
\item 
  About me:
  \begin{itemize}
  \item 
    I host a bi-weekly oggcast, Free as In Freedom: htp://faif.us
  \item 
    I'm on identi.ca: @bkuhn
  \item 
    I have a blog: http://ebb.org/blog
  \end{itemize}
\end{itemize}
Copyright © 2008, 2009, 2010, 2011 Bradley M. Kuhn. These slides,
this talk, and audio/video recordings thereof are under a Creative
Commons Attribution-Share Alike 3.0 United States License.

\end{document}
